\documentclass[bfivepaper,twosided,justified,nobib]{tufte-book}

\hypersetup{colorlinks}% uncomment this line if you prefer colored hyperlinks (e.g., for onscreen viewing)

%%
% Book metadata
\title{How I Learned to \\ Worry Productively about AI: \\  \\ Robot Apocalypses \\ Averted}
\author{Z. Irene Ying and Jordan Boyd-Graber}
\publisher{University of Maryland}

%%
% If they're installed, use Bergamo and Chantilly from www.fontsite.com.
% They're clones of Bembo and Gill Sans, respectively.
%\IfFileExists{bergamo.sty}{\usepackage[osf]{bergamo}}{}% Bembo
%\IfFileExists{chantill.sty}{\usepackage{chantill}}{}% Gill Sans

%\usepackage{microtype}

\usepackage{booktabs}
\usepackage{graphicx}
\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
\graphicspath{{figures/}}

% The fancyvrb package lets us customize the formatting of verbatim
% environments.  We use a slightly smaller font.
\usepackage{fancyvrb}
\fvset{fontsize=\normalsize}

%%
% Prints a trailing space in a smart way.
\usepackage{xspace}

\newif\ifcomment\commentfalse
\input{preamble}

% Generates the index
\usepackage{makeidx}
\usepackage{natbib}
\setcitestyle{authoryear}

\makeindex
\newcommand{\blankpage}{\newpage\hbox{}\thispagestyle{empty}\newpage}
\begin{document}

% Front matter

\frontmatter

% \tableofcontents

\input{characters}
\input{chapters/00-frontmatter}
\input{chapters/10-intro}
\input{chapters/20-story1}
\input{chapters/25-cyberjerks}

%\chapter{General AI}
%\label{chap:gai}

%\input{chapters/example}

\newcommand{\outline}[3]{\chapter{#1}
  \begin{itemize}
  \item {\bf Story:} #2
  \item {\bf Essay:} #3
  \end{itemize}
}

\outline{General AI}{\lastname{} is called in to consult on the
  deployment of a general AI that a military contractor has developed.
  It breaks out of its sandbox, causing physical havoc with military
  equipment.  A specialized AI designed to contain the general AI is
  surgically deployed to resolve the problem.}{This chapter explores
  the distinction between general and specialized
  AI---\specializedai{}---and how the pressures and history of a
  general AI cause them to be vulnerable to specialized algorithms and
  methods.

  With respect to current day, the chapter reviews the approaches
  toward general AI (and how they haven't gotten anywhere) as opposed
  to specialized AI.  }

\outline{Resource-Constrained AI}{\lastname{} is comfortably faculty
  at University when her past history with \specializedai{} comes back
  to haunt her: a rogue actor has found an old copy and using it to
  terrorize a small Canadian town.  \lastname{} works with the local
  authorities to deploy their specialized AI to counteract the
  threat.}{This chapter explores how governments' role as a protector
  of populations evolves in a world with ubiquitous and powerful
  artifcial intelligence.  Although in some ways AI is democratizing,
  it is still connected to access to real-world resources (material,
  energy, physical space), and governments and multinational
  corporations will have the most powerful and capable AI agents.
  Just as we trust governments with technology that can end the world
  (nuclear weapons), we must trust the government to responsibly
  harness the power of artificial intelligence.

  This discussion will be connected to the connection to electricity
  and specialized hardware for bitcoining mining and deep learning.  }

\outline{AI's Role in Social Interactions}{\lastname{} finds herself
  drawn into a political scandal when an AI agent begins influencing
  public opinion against her and her research.  She fights back the
  only way she knows how: with AI.}{While the threat of AI is often
  seen as physical, another salient fear is social and emotional: AIs
  can play on fear and emotion.  In many ways this is harder to notice
  and fight against because these effects might be superficially
  welcome by their targets.  This chapter talks about how AIs can
  detect, influence, and interact with people.

  Examples will be drawn from AI approaches to detecting deception and
  manipulating audiences on social media.  }

\jbgcomment{Not putting in outline, but what might be a way of making
  the plot work, this could be a scheme by a faculty adversary that
  she uncovers, which leads her to sell out.}

\outline{Cashing Out}{Fed up by politics and the politics of academia,
  \lastname{} goes off to make a pile of money at \fintech{}, turning
  her AI smarts to Wall Street.  It's harder than she
  thinks.}{Financial markets, like much of AI, hope to make
  predictions about what will happen in the world and to estimate
  uncertainty.  This chapter investigates whether capitalist markets
  are possible in a world with effective AI.
}

\jbgcomment{Connection between the two chapters: previous chapter allows everyone to go out an earn a capitalist wage.}

\outline{The Silent Revolution}{Having saved the world multiple times,
  \lastname{} retires to spend more time with her grandchildren and
  sees their relationship with AI.}{The book concludes with how AI can
  change social interactions and society in less overtly negative
  ways, warning about a future where AI can diminish the richness of
  the human experience (if we let it).

  More of an opinion piece, this chapter focuses on how AI and humans
  should interact with each other.  Specifically, arguing for
  cooperation and collaboration rather than outright replacement.  }

\backmatter

\bibliography{bib/jbg}
\bibliographystyle{plainnat}


\printindex

\end{document}
