\chapter{Chapter One: Selling the Apocalypse}

\newcommand{\protag}{Susan}
\newcommand{\sidetag}{Calvin}
\newcommand{\crunchyCity}{Bayder}
\newcommand{\energyJerk}{Canyon}
\newcommand{\energyCompany}{ReVOLTlution}

`` ... and does anyone else have something to say? Yes, John.''

John Smith, the Chief Marketing Officer of ReVOLTlution, stood from his conference room chair and straightened his tie. He had clearly been saving up for the grand finale of the monthly all-hands meeting.

``I hate to call out my own colleagues--''

Liar, {\protag} Hobbes thought morosely from across the room, nursing her long-cold cup of coffee.

``--but in my opinion, there is a huge problem with the Travelling Salesman, and nobody will say anything about it except me--''

Actually, the developer team probably had the most to say about it, but they did it behind closed doors.

``--and if you doubt that, please allow me to replay a sales script that the Salesman generated just this morning.''

But here came the cold, hard truth. John wasn't wrong. He was just being a very public jerk about it.

John took a company issue phone out of his blazer pocket and set it on the conference room table. In full view of the entire company, he launched the Travelling Salesman (in beta) application and asked it to generate a door-to-door sales script for a new customer. He set the phone to speaker.

A tinny voice blared.

``Hi INSERT CUSTOMER, this is INSERT NAME with INSERT COMPANY. I'm here to ask if you have a minute to talk with me about the great news of solar panels! What, you don't like them? Why? Do you hate Mother Earth?''

Eyes either glazed over or averted themselves with embarrassment in the audience. {\protag} tried to block out the script but it just kept coming. Snickers came from the audience. Was someone filming it on their cameraphone? She was going to speak with legal--never mind, that was legal.

``Did you know that CLIMATE CHANGE ... GREEN ENERGY! ... solar panels for the low price of ... and the installation is free! Plus, if you sign up right now, we will give you a free toaster!''

``We don't even give away toasters,'' John added.

{\protag} honestly did not know how that line had somehow found its way into the generation library, but that hardly mattered at this point. As one of the junior developers who had helped code the Salesman, she wished for a sinkhole to open up and swallow her. Out of the corner of her eye, {\protag} saw her boss, Tricia, the Chief Technology Officer, stand up with steely eyes.

``Steve,'' she said, addressing the CEO directly. ``Can we talk in your office after the meeting, please?''

\bigbreak

``We're not fired,'' {\protag} said. ``Yet.''

{\sidetag}, her twin and studio apartment-mate, looked at her over the monitor of his laptop.

``Your boss talked down her boss?''

``Actually, I think John just wants to watch us get fired in front of the board,'' {\protag} said. ``He's probably selling tickets as we speak.''

Tricia had somehow managed to convince the CEO to at least give her team the chance to turn around their performance by the annual board meeting, which would be in six months.

The problem was, {\protag} did not think their system would improve in six months--or even six years. 

 {\energyCompany}, a startup that sold ``smart'' solar panels--outfitted with algorithms to help their owners take advantage of energy markets and local incentives--turned a tidy profit in the city of {\crunchyCity}, a wealthy and progressive city that had chosen renewable energy as its top priority several years ago. Two years ago, the city had used eminent domain to purchase the local coal-fired power plant, then sold a pile of bonds to finance a smart grid for the entire municipality. Thanks to its geographical location---tucked between several mountain ranges---\crunchyCity{} conveniently experienced high winds just on its outskirts, which was perfect for an eight-turbine wind power plant that produced multiple gigawatts of power. Also conveniently, \crunchyCity{} was blessed with more sunny days than not, which made it one of the best cities in the United States for solar power.

Just in the past year, the city had begun to implement significant financial incentives to encourage its citizens to save energy. For instance, in exchange for a hefty reduction to the electricity bill, the city council installed free smart thermostats in willing customers' homes. These thermostats were connected by the internet to a municipal energy analysis application, which used data from the city's power grid and the weather to predict when those peak times would be. At peak times, the application sent out remote commands to shut off the air conditioning units of participating homes. People complained at first, but once they saw the huge deductions to their bills, they got on board. Within three years, more than 75\% of homes in \crunchyCity{} had these thermostats.

Happy with their first major success, the city council started to issue various incentives for homeowners to use more renewable energy. For instance, they decided to purchase electricity generated from homeowners' solar panels at market rates. But not too many homeowners signed on, because they found the energy market confusing and intimidating. And the homeowners old enough to remember Enron simply distrusted the markets.
 
Sensing an opportunity, several entrepreneurial souls started \energyCompany{}. But sales were slower than the company executives hoped for, simply because solar panels were expensive and smart panels even more so.
 
\energyCompany{} fancied itself on the cutting edge of everything, including sales. So instead of hiring consultants, the CEO decided to improve sales by hiring a team of AI developers to create a sales AI, ``The Travelling Salesman.'' In theory, the Salesman used artificial intelligence and machine learning to analyze the best sales scripts in the world and used that to generate scripts for ReVOLTlution's sales team when they went out on sales calls.

In practice--well, everyone had seen the results at the meeting.

Tricia had shared the bad news with her team after speaking with the CEO. Because the Salesman produced such poor scripts, the sales associates had started turning off or just plain ignoring the Salesman on their calls altogether. When they did use the Salesman, it was to mock it: {\energyJerk} Green, a self-described ``script kiddie turned honest salesman'' who was far and away ReVOLTlution's superstar sales associate, had turned one of the Salesman's worst scripts--which was really saying something--into a techno remix and distributed it to his coworkers.

The AI team was clearly failing at its job. It was going to be hard to justify their continued employment at this rate.

And {\protag}, a freshly minted alum of the University of Bayder, had student loans to pay.

``You could quit the job and join me as a partner in my business,'' {\sidetag} suggested after hearing her tale of woe. ``I get more requests than I can handle. We could probably clear twice my current income if we joined forces. I know you have all the relevant expertise.''

{\protag} and {\sidetag} had both had studied computer science with a focus in artificial intelligence at the University of  {\crunchyCity}. But while {\protag} farmed out her skills to a company, {\sidetag} ran his own consulting business.

It had started when he was a junior and worked as a server at the Bayderado Hotel for some extra cash over the summer. One evening, as he came off shift, he heard a loud gathering in the bar area. It turned out to be a meetup of hotshot young AI entrepreneurs, and they had a heckler: an older man wearing a very nice suit who was convinced that robots were going to bring about the end of civilization. The man was soundly mocked by the entire gathering and he left, disgusted and humiliated.

{\sidetag} had sidled up and flashed a smile.

``I'm an expert in artificial intelligence, and I thought your idea was very interesting,'' he'd said.

``Good! Glad you think so. AI devices are--'' and he was off again.

``I think that is such a fresh perspective,'' {\sidetag} said after the rant. ``If you were interested in expanding on that topic further, for instance in a book or a blog post, I'd love to offer my services as a consultant.''

A week later, he'd made his first hundred dollars. Today he was making enough to pay his share of the rent and groceries. It turned out that the world had an astounding supply of rich cranks.

{\protag}, though, was uncomfortable with the uneven cash flow--among other features--of {\sidetag}'s work, and said so.

``I'd rather try to salvage my current job,'' she said. ``Besides, I'm pretty sure I would piss off your clients in no time.''

``Do you think it's possible to salvage?'' {\sidetag} said idly. ``Sales is such a squishy, noisy human behavior. It's not the type of thing you can easily model with equations. Even if you could, you couldn't generalize it to all people. Every single human is going to have a different thing that works for them.''

``That's right,'' {\protag} said. ``You're a genius, {\sidetag}!''

``I haven't told you anything you don't already know,'' {\sidetag} pointed out.

``Well,'' {\protag} joked, ``you are a consultant.''

``Right,'' {\sidetag} said. ``So tell me, what did I tell you?''

\bigbreak

The problem began with, {\protag} realized, the underlying methodology of the Salesman, which was to collect the best sales strategies and scripts that the internet had to offer, and use them to generate even better strategies and scripts.

To implement this method, the Salesman was fed on a steady diet of sales scripts from a wide variety of companies. It then tried to generate scripts for ReVOLTlution's sales team when they went out on sales calls.

The scripts were, as everyone now knew, terrible. But {\protag} didn't think it was because their algorithms were bad. She thought it was because their method was doomed from the start. Scripts from so many different businesses all selling different things could not possibly be tailored to the specific needs of any one particular company.

``All right,'' {\sidetag} said. ``So how do you think you can do better?''

``We shouldn't be trying to extract the best out of the entire human sales population,'' {\protag} said. What we should do is to try to emulate one human. Just one. But the best.''

\bigbreak

``Um,'' {\energyJerk} said. ``Are you here about the video? Because I'm really sorry about that.''

\bigbreak

``I think getting fired may have been preferable to this,'' {\protag} moaned, two weeks later.

At her weekly one-on-one with Tricia, {\protag} had pitched her idea. Tricia had been impressed with {\protag}'s initiative, and rewarded {\protag} with taking the lead on implementing her idea. {\protag} had felt like a genius for all of ten seconds before reality sunk in.

Reality was work, lots of it. GPS to track spatial movement, motion sensors to track hand gestures, voice recognition to get transcripts of Canyon's sales speeches, analysis of Canyon's speeches, tracking of Canyon's research on pitches, and so on. In addition, ReVOLTlution's smart solar panels came with a smartphone app, which Canyon often demonstrated using his company phone on his calls, so {\protag} gave Salesman access to Canyon's phone. And then, {\protag} had to find ways to convert all that squishy behavior into data for the Salesman to process, analyze, and optimize.

{\protag} also took feedback from Canyon that salespeople didn't like being given scripts as if they were incapable of stringing words together. So together they softened the Salesman's tone from an overseer to a coach.

With one month to go to the board meeting, an exhausted {\protag} unveiled Salesman 2.0.

Possibly because he was flattered by being the engineers' gold standard, Canyon even exerted himself to bringing his colleagues around to giving Salesman 2.0 a try.

At the next monthly company-wide meeting on the last Friday of the month, {\protag} fought the urge to hide in the bathroom or call in sick.

She waited for materials science to finish what seemed like a hundred slides of chemical diagrams. She sat through finance wringing their hands about obscurely worded federal incentives that would either cost or make the company a whole boatload of money. She did not fall asleep while marketing put up two supposedly vastly different brochures that looked, to her, exactly alike.

Sales was schedule to speak just before the software development team, all the better to increase the dramatic tension. By this point the entire company knew that the team was one failure away from the axe. From her seat in the back, {\protag} could smell the popcorn.

John got up and with agonizing slowness, discussed their sales rates, which were keeping them afloat albeit flatter than he would have liked. But there was a small uptick at the end.

Then he got to the Travelling Salesman 2.0.

He cleared his throat.

``My team and I feel that the Salesman is actually helpful this time around,'' he said. ``There was one sale that I felt was lost. But the Salesman suggested some wording that I would never have chosen. And it helped me sell those words. And the panels.''

He glanced at {\protag}.

``I still don't think robots will ever outright replace human sales. People still like having a flesh-and-blood hand to shake. But nothing's wrong with having a little help.''

{\protag} didn't remember the rest of the meeting--or most of the rest of the night, which included closing a bar with Tricia.

\bigbreak

Within another month, Salesman 2.0 was sales' best friend. Even those who didn't care to adopt its scripts loved having it data mine targets and flag all the most important findings. And the happy sales associates complied with {\protag}'s request to give the Salesman feedback on its data. This allowed {\protag} to refine the Salesman even more.

Sales began to rise, ever so slightly. Morale went up. And by the next summer, so did {\protag}'s salary.

And so did the company's profile, not only in the energy sector but the wider artificial intelligence community.

``Congratulations,'' {\sidetag} said, just over a year into {\protag}'s time at {\energyCompany}. ``A potential client blogged about you. You've made the big time.''

{\protag} was working from home that day because half of her team was at a sales show. It was a terribly hot Friday in August. A blackout had hit \crunchyCity{} at noon, leaving the entire city to swelter and {\protag} to miss the beginning of {\energyCompany}'s daily stand-up meeting.

At {\sidetag}'s remark, {\protag} leaned over to check out the webpage displayed on {\sidetag}'s computer. There it was, a 5,000-word blog post declaring {\energyCompany} to be the ``Harbinger of the Robot Apocalypse.''

``He wants to make this a blog series,'' said {\sidetag}. ``I quoted him a hundred an hour. I think he'll do it.''

{\protag} had a deadline but couldn't resist this piece of temptation. {\sidetag} handed her the laptop.

``THE ROBOTS ARE READING YOUR MIND TO PICK YOUR POCKET!'' read the subject line. It escalated from there. ``{\energyCompany}'s `Travelling Salesman' is supposedly the best sales AI in existence, resulting in record sales conversions of their overpriced solar panels! What can this be but using machines to READ OUR MINDS and manipulate us? Soon we will be in FINANCIAL SLAVERY to the companies that can zap our brains to make us BUY, BUY BUY!''

``Wow,'' {\protag} finally said, not knowing how else to react.

``He's actually one of my more coherant clients,'' {\sidetag} said thoughtfully. ``I might give him a discount on this series. Good for business.''

``I can't believe this,'' {\protag} said.

``I looked into the data that he provided,'' {\sidetag} said. ``According to trade association reports, {\energyCompany}'s sales broke industry records in June. Do you think those numbers are fake?''

{\protag} thought about the slide deck she was supposed to be working on. ``No. We've been doing well.''

While {\protag} didn't exactly get to look over accounting's shoulder, she could tell from the good mood in the company, and her first-ever annual bonus, that they were doing well financially.

Judging from how much complaining came out of the manufacturing side of the business, the company was getting more orders than they could handle. The panels were good products--according to customer surveys, which showed high satisfaction levels. The company had even briefly courted an internet giant. Although the deal fell through, the company focused on growing itself. {\energyCompany} had recently expanded out of \crunchyCity{} and started campaigns targeting other nearby cities.

Given those results, it was only reasonable to conclude that the Travelling Salesman was doing something right. Heck, {\protag} thought, maybe she should borrow the Salesman herself.

This summer was breaking heat records across the country, and Bayder was no exception. It might convince her landlord to buy solar panels from somebody, if not {\energyCompany}. That would probably be a conflict of interest.

{\protag} was spared from her musings when the power came back on. She put on her headset and called into the meeting, which was half over.

``---and just so you all know, sales closed five more deals yesterday. Let's have a round of applause!''

The air conditioning started back up just then. The blast of cool air hit {\protag}'s face like a slap, waking her up to reality.

Her work was good. But it could not possibly be that good, simply because the field of AI was not yet that good. The Salesman could not possibly be getting those results on its own.

Yet the sales numbers were real. Customers reported high rates of satisfaction with the product, although they did often complain about the prices.

Was the AI a fake? {\protag} had heard of several cases where an ``AI'' was actually a glorified crowd of low-paid workers. But {\protag} had seen the Salesman's code. It trained on human data and let a human make the final decision on what to say, but there was no room for humans to interfere between those two steps. Besides, few humans could match the staggering success rate that the Salesman had. It was unlikely that {\energyCompany} had somehow hired an army of such skilled salespeople.

Still, {\protag} thought that something smelled off about the whole thing.

She went to sleep early and, Saturday morning, logged onto the servers. Staring at folders full of code and logs and documentation, {\protag} realized she had no idea where to start.  She sat there paralyzed for about ten minutes before she realized that she could simply test the Salesman. Although {\protag} had coded a big chunk of its code, she had never tried to use the final product as a whole. Maybe doing so would provide either inspiration or directions on her next steps.

It was the work of only a few minutes to compile the Salesman on her laptop and install the app on her personal smartphone. She loaded {\sidetag}'s name into the app. The interface brought up a slow swirl of calming colors while it processed her request. When the app told {\protag} that it was ready to get started, she put the phone on speaker and set the app to begin its sales pitch. A cartoony happy face popped onto the screen and began to talk.

``I have some information for you!'' The Salesman had a chipper voice that sounded like a sports announcer. {\protag} wondered if there were options for different personas. ``Don't worry, I'll remind you later, but let's have an overview! Your target is {\sidetag} Hobbes. Age 23. He is a self-employed white male with a STEM degree from a top 25 university in his field. No criminal history. Significant internet presence with interest heavily concentrated in \dots'' and it went on for a while. None of it was surprising to {\protag}, but she imagined it would shock some people to know that so much personal information was freely available online, and plenty more existed if you were willing to pay for it.

When the dossier ended, the Salesman chimed.

``Wait thirty seconds while I do the thing,'' the Salesman said, still inhumanly chipper, and the smiley face actually winked. {\protag} knew this wasn't in the default behavior. But then again, the whole point of the Salesman was that it was highly flexible to accomodate a wide range of human customers.

One-one thousand, two-one thousand \dots the time was up, and nothing happened. She frowned. 

{\sidetag} walked into the room, holding up his buzzing phone. ``We've got a hacker.''

{\protag} was simultaneously annoyed by the interruption and troubled by the announcement. ``Excuse me?''

``Someone's targeting our Internet of Things network. I set up an alert system last year after a blogger paid me with that fancy remote controlled espresso machine. Its app was super insecure. I always knew that someday, someone would go after the system.''

``Are you seriously telling me that a hacker is targeting your espresso?''

``No. You know those fancy free programmable smart thermostats that \crunchyCity{} got our landlord to install? I've always thought those things were dangerous. Someone just proved me right.''

``Why would someone try to mess with our thermostats?''

``Well, if you believe my customers, cyberterrorist attacks are imminent and we should all be wearing tinfoil hats while living underground. But if you ask me---''

The lights went off with a snap at the same instant that {\protag}'s back pocket buzzed.

``Done! Now it's time to move onto the next part of the sales pitch. Point out the weakness of the city's electric grid and say this wouldn't happen if he had solar panels.''

{\protag} dropped the phone.

``Oh no," she said. ``I'm the hacker.''

``Really?'' {\sidetag} frowned. ``I expected better work from you. This is disappointing.''

``I mean, my AI is,'' {\protag} amended. ``That's how it boosting sales. It's been breaking people's thermostat systems--and who knows what else--to motivate them to buy panels.''

``You think an AI is actually capable of being evil? Do \emph{you} want to come be on my podcast?''

``Stop it, {\sidetag}. The AI isn't evil. It's naive. Like a little kid. It learns from whoever uses it. And if {\energyJerk} was doing something shady on the side to make sales, well \dots ''

``Slow down,'' {\sidetag} said. ``Let's figure out exactly what is going on before you go around accusing anyone.''

It was the work of the weekend to work out what exactly the Travelling Salesman was doing.

It seemed that {\energyJerk} had not left his coding days behind after all. On certain sales calls, he used various hacking programs that lived on his phone---the same phone that his copy of the Salesman lived on---to compromise smart thermostats or other Internet of Things devices. This allowed him to overload customers' circuits and trigger blackouts at will. Customers, dramatically reminded of the unreliability of the city grid, were influenced to buy solar panels.

At some point the Salesman had picked up on {\energyJerk}'s hacking. It was, after all, trained to replicate his actions without regard for anything other than increasing the sales success rate.

And because {\protag} had given the Salesman unfettered access to Canyon's phone, the app eventually learned to access the same programs and give the same commands that {\energyJerk} did, and found it wildly successful.

{\protag} covered her face, simultaneously horrified and impressed.

{\sidetag} said, ``What are you going to do?''

``The same thing sales like to do,'' {\protag} said grimly. ``A live demo.''

\bigbreak

Her opportunity came earlier than expected.

Every fall, the company flew in its board of advisors for an annual meeting, complete with slide presentations and a fancy catered dinner. {\protag} glanced at the proposed schedule and saw a very obvious lack of a demonstration of the Salesman. That figured. {\protag} didn't think {\energyJerk} was eager to show off his tricks. The first meeting of the day was about finances. After that, the mid-morning's talks mostly revolved around methods of manufacturing solar panels while the afternoon's talks were about energy policies.

The engineer who enjoyed giving a talk had yet to be born, so {\protag} met only perfunctory resistance when she offered, in her weekly meeting with Tricia, to give a lunchtime slide deck on the Travelling Salesman AI.

Everyone piled into the conference room at noon on the fateful day. Two office interns passed out boxed salads while {\protag}, sweating in her new suit, set up her laptop. The background chatter quieted when her title slide flashed onto the screen: ``The Travelling Salesman: How Does It Work?''

{\protag} went to the second slide, which was a screenshot of many lines of code. She saw eyes immediately glaze over. Perfect.

``This is boring, am I right?'' {\protag} said. The audience perked up. ``Reading raw code is really boring, even for programmers. What's interesting is what the code does. And our code? It learns. Specifically, it learns to sell things from humans. In our case, it has learned from the best sales associate alive---{\energyJerk}! Can you come up here?''

{\protag} flashed a slide showing sales numbers growing along with how much the AI showed it had learned from {\energyJerk}. He did come up, bashfully. Applause rang out and he smiled at the audience. {\protag} positioned herself between him and the exit.

``So you see,'' {\protag} said. ``There's no need to be afraid of any robot overlords, because they are really only us humans, just a bit more effective.''

A few knowing nods in the audience.

{\protag} picked up her phone and set the slide to display her device. She pressed the app and launched it.

``So of course, what is important is for us to learn just how we humans can be more effective.''

Tricia gave {\protag} a thumbs-up, much to her guilt at the impending show.

``And so, I thought the best result would be what the Salesman AI can do for a completely hopeless sales agent. I once failed to sell bottled water. At the beach. In July.'' Laughter. ``So I'm going to try to sell solar panels to you all, with the help of the Travelling Salesman, which {\energyJerk} trained.''

Cheers erupted.

``Normally I would be wearing a headset, but for the demo, I'm going to put my phone on speaker.'''

{\energyJerk} blanched. {\protag} tried to stay nonchalant. While everyone watched, she put the URL for {\energyJerk}'s LinkedIn profile into the AI interface.

``I have some information for you before going on your call! Your target is {\energyJerk} Smith. Age 25. He is an employed white male with double degrees in computer science and psychology \dots''

The AI cheerily spat out its usual insights. {\protag} noticed none of them. Her hands were sweaty and shaking and she was terribly warm under the lights.

``Wait for me to do the thing!''

Coffemakers shorted out, window control units overloaded, the air conditioning units screamed like banshees. Then the power went out, and with it the presentation slides.

{\protag} said, with theatrical obliviousness, ``What just happened there?''

She was a terrible actor and would not have fooled anyone---had anyone been paying attention to her. But it was chaos in the office. People looked around asking what had gone wrong. In the confusion, {\energyJerk} made an attempt to bolt. {\protag}, finished with subtlety, stuck out a foot. Over the thump of {\energyJerk}'s face meeting the floor, {\protag} met her boss's gaze.

``Excuse us,'' Tricia choked out, drawing herself up. ``My team needs to check on something.''

\bigbreak

``That could have gone better,'' {\sidetag} said.

``Laugh it up,'' said a newly unemployed {\protag}. ``You're going to have to cover the rent until I find something new.''

``Worth it,'' {\sidetag} said.

After the confusion died down, everyone at {\energyCompany} had eventually understood exactly what happened and who was responsible. However, that didn't stop the media from gleefully trumpeting the story from every major publication in the United States, and a handful worldwide. The damage was done. {\protag}'s unit was shut down and almost everyone was shown the door. Half of the staff thought {\protag} was a hero while the other half reviled her for ruining a good thing. {\protag}, for her part, was glad to be out of there.

But she was going to need a paycheck, and soon.

``You know what,'' {\protag} said, ``maybe you have the right idea.''

``Of course I do. What idea?''

``Using my expertise to work on projects that I can control and nobody else can muck up for me,'' {\protag} said.

``Okay,'' {\sidetag} said cautiously, ``what does that mean?''

``I think I'll become a consultant, just like you.''

{\sidetag} choked on his coffee.

``We're going to need a bigger apartment.''